Improving \textsl{Gaia} parallaxes with a data-driven model \\ of the color--magnitude diagram

### 2017-06-02 paper/paper.tex
discussion rearrange

\BLcomment{This whole section is highlighted because I have moved paragraphs around and also added small clauses here and there. I've tried to improve the logical motivation and presentation using the (great) paragraphs you wrote. Feel free to revert if you don't like it.}\BLedit{

Under these assumptions, we would like to infer the true parallax, $\varpi_{\true}$, for each star given its measured parallax, $\varpi$, and associated uncertainty, $\sigma_{\varpi}$, from \tgas.
A first application of Bayes theorem leads to
\begin{equation}
p(\varpi_{\true} \given \varpi, \sigma^2_{\varpi}) = \frac{1}{Z}\,p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi}) \, p(\varpi_{\true}) \quad ,
\label{eq:bayes}
\end{equation}
where $p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi})$ is the likelihood of the observed parallax, $p(\varpi_{\true})$ is a prior pdf for the true
parallax, and $Z$ is a normalization constant \BLedit{(the model evidence, or marginalized likelihood)}.
The strength of our inference lies in how we generate our prior.
Instead of assuming a functional form for the spatial distribution of stars in the Milky Way (\citealt{astraatmadja16b}), or relying on stellar models (\citealt{gaia16}), we build a noise-deconvolved estimate of the \cmd\ from the observed colors and absolute magnitudes of our full set of stars.
This is because supplementary to the trigonometric measure of distance, the photometry of stars is also informative of their distances.
The intrinsic temperatures and luminosities of stars tightly correlate leading to the observed colors and absolute magnitudes of stars also tightly correlating, usually visualized as the \cmd.
The tight correlation of the \cmd\ benefits our inference because the luminosity or absolute magnitude of a star $M$, when compared with its brightness or apparent magnitude $m$, contains distance information $d$ via the fundamental relation $m=M+5\log(d/10\, \mathrm{pc})$.
We include this in the inference by using the \cmd\ as part of our prior for $\varpi_{\true}$.
Specifically, we use the \cmd\ generated from the data themselves, as detailed in the next section.

@@

We can now write the multidimensional analogous to Equation \ref{eq:bayes},
\begin{equation}
p(\mathbf{Y}_n \given \mathbf{y}_n, I_n) = p(\mathbf{y}_n \given \mathbf{Y}_n, I_n) \, p(\mathbf{Y}_n) \label{eq:posterior}
\end{equation}
where $p(\mathbf{y}_n \given \mathbf{Y}_n, I_n)$ is the likelihood of the data vector, and $p(\mathbf{Y}_n)$ is our prior.
Before we describe how this can be turned into a posterior belief on the true parallax, we turn our attention to the prior, which will be generated from the data themselves. We will use XD, which will take advantage of the Gaussian likelihood.
}

@@

\BLedit{To infer $\varpi_{\true}$ once the \cmd\ prior is constructed, we first do the inference in the 2D space of the \cmd\ prior $p(\mathbf{Y} \given \alpha)$, using Equation \ref{eq:posterior}.} We then project the 2D posterior into the 1D transformed absolute magnitude space, and do a little algebra to solve for $\varpi_{\true}$

@@

\BLcomment{Needs to say much more - unfortunately I cannot complete this part since I don't know what you've done in enough detail.}

@@

\BLcomment{Isn't this repeating what was described in the previous paragraph?}
We need to correct the photometry used in our method, both for the generation of the prior and the likelihood, for dust. With the emergence of 3D dust maps, it is now possible to estimate dust corrections to stars within the Milky Way. The challenge is getting a proper estimate of the distance to the star before inferring $\varpi_{\true}$, since we need a dust correction prior to doing our inference. Dust corrections are particularly non trivial for stars with poor parallax signal to noise. Stars with poor signal to noise have a likelihood that is consistent with infinite distance and can therefore have severe dust corrections. This is most obvious in the giant stars, which tend to have the lowest signal to noise, and can therefore have dust corrections $> 2$ magnitudes. Although the dust model is probabilistic, as well as our distance inference, to move forward, we break from a fully Bayesian framework, in which we might infer the distance and the dust simultaneously (more about this in the discussion). Instead, we correct for dust using a point estimate from the 3D dust map \citep{green15} sampled at a distance estimated from iteratively inferring and then sampling our posterior pdf over $\varpi_{\true}$. In more detail, to apply a dust correction to the \tmass\ photometry, we first generate our prior using the observed, attenuated photometry. Using this raw prior, we infer more precise parallaxes to all the stars in our sample. We then use this more precise parallax posterior to get a measurement of dust in the 3D dust map for each star. And repeat. This approach minimizes severe distances, and their associated severe dust corrections, which would occur if using the likelihoods alone. In detail, we take the $5\%$ quantile of each distance posterior, so the closest bit of the posterior, and sample the 3D dust map at that distance. We take the $5\%$ quantile, as apposed to the median, for example, to do a light dust correction, because over correcting for dust can shoot the likelihood into an unphysical, different space of the \cmd\ and bias the parallax inference. For each distance, we sample the median of the probabilistic dust model. We apply each light dust correction to our \tmass\ photometry given by Equation \ref{eq:data} where $Q_{\lambda}$ = [0.709, 0.302] for bands [J, K] respectively, taken from table (REF TABLE SFD). We then regenerate our prior using the lightly dust corrected photometry. The dust values we obtain in our final iteration, and which we apply to our photometry, are visualized in galactic coordinates in Figure \ref{fig:dust}. After applying a dust correction, we do not update the uncertainties, but keep them the same as before the dust correction, which is wrong and under representing our uncertainties but this leads, again, to a lighter deconvolution. \BLedit{In other words, the features of our inferred \cmd\ are conservatively broad (broader than if we over-estimated the errors which would have lead to excessive deconvolution and unrealistically narrow features).} We could add the covariance of the dust to our likelihoods, but this is beyond the scope of this demonstration of concept. We iterate this process 10 times, however the dust values seem to converge after about 2-3 iterations. The iterative process allows for slight variations in distance and corresponding dust to become more consistent with the \cmd.

@@

\BLedit{
\subsection{Shrinkage of parallax uncertainties}

@@

}
\BLcomment{Need to discuss the implications of that.}

@@

Transition clause/sentence. We obtained the photometric parallaxes in this project
by building a data-driven model of \cmd\ of the stars in the \tgas\ data set,
and using it as a prior pdf for Bayesian inference.
The posterior pdfs for distance that we obtain are, in general, much
narrower than the likelihood functions delivered by the
\gaia\ Mission, and therefore the distance estimates (or,
equivalently, parallax estimates) are much more precise.
It is not surprising that a Bayesian inference provides more
precise inferences than the likelihood function alone; Bayesian
inferences bring in new information that decrease variance (but
often introduce bias).
We have shown that, in addition to pure precision improvements, at
least some aspects of \emph{accuracy} have been improved as well, by
showing that the posterior distance estimates to stars in (at least one) stellar
cluster are much more clustered than likelihood-based distance
estimates.

Transition clause/sentence. Although we have not performed principled hierarchical Bayesian inference in this
project, we built our prior for each star's parallax by
performing a statistically responsible deconvolution
of the \gaia\ \tgas\ data that is justifiable under a clear set of
assumptions.
This deconvolution, using a reasonable noise model and an
assumption of stationarity, shrinks the parallax uncertainties, most dramatically for the stars measured at lower signal-to-noise.
Because the method we use, \xd, which is
an Empirical Bayesian maximum-marginalized-likelihood estimato, accounts for
heteroskedastic noise, we were able to build our photometric parallax
model---which is a model of the \cmd---using
all the data, not just the data with the highest signal-to-noise.
So, the model for the \cmd\ we built is representative for
our subset of the \tgas\ Catalog.
Because our model is representitive of our full subset catalog, we expect any (sensible) parallax estimates
we generate from our parallax posterior pdfs to be only weakly biased.

Transition clause/sentence. The empirical Bayesian methodology is a good approximation to full
hierarchical inference in the limit of large numbers of data points, such that no individual star is carrying a lot of weight in the
model of the \cmd.
With its millions of stars, the \tgas\ Catalog appears to safely
live in this limit.
However, because there are parts of the \cmd\ that are poorly populated,
even in \tgas, the empirical Bayes approximation may be bad for some
portions of the \cmd.
In particular, there appear to be essentially no white dwarf stars,
and other classes (like bright red giant branch stars) have few to no exemplars
in the data set at high signal-to-noise (in parallax).
For all these reasons, we do think that this approximation is not perfectly
safe, and it is a goal to explore full hierarchical modeling in the coming
years (and see, for example, \citealt{leistedtHogg2017}).

Transition clause/sentence. The biggest limitation of our method for generating the \cmd\ from the data is that we made no use of any kind of \tgas\ selection
function or completeness estimate or inverse-selection-volume corrections.
This model is a model of the \tgas\ Catalog
(or really the \tgas--\tmass\ intersection), not of the stars in
the Milky Way, nor any volume-limited subsample of the Milky Way.
Importantly, because we used \emph{all} of the stars in the \tgas--\tmass\
intersection, the model \emph{is} representive of the full intersection,
even the distant parts, where all the parallaxes are measured at low
signal-to-noise.
This is in contrast to techniques that might build the model from only
the high signal-to-noise sources; such a model would be biased towards
stellar types and compositions found locally, as well as lower-luminosity stars,
and stars which happened to get good parallaxes.
However, our use of the entire \tgas--\tmass\ intersection without
any selection function restricts the use of our \cmd\ model as a prior
to inferences within that same \tgas--\tmass\ intersection.
It would be biasing to use this same prior to infer
distances for the full billion-star catalog released in \gaia\ DR1
alongside \tgas, or for stars fainter or brighter or bluer than those
included in the \tgas--\tmass\ intersection.
\BLcomment{I would specifically mention again the assumption of stationarity, and explain how it is broken by the selection function (in particular fluctuations in the detection probability of objects as a function of position, parallax, photometry, etc). Here, or in the critical discussion of assumptions.}

Transition clause/statement. The outputs of this project include posterior pdfs for stellar parallaxes.
It is tempting to treat these outputs as equivalent to some kind of catalog of
measurements, as we treated the original parallax measurements from \tgas.
That is, it is tempting to treat these as simply ``better'' measurements.
Taken one star at a time, this is permitted and true, and the basis for our
claim that we have de-noised the \tgas\ Catalog.
However, there are important differences, as there are with all
probabilistic catalogs (\citealt{hogg11}; \citealt{portillo17}), between the
measurements and the de-noised measurements.
For users of our output, the most important difference between our
output and the \tgas\ data input is that the \tgas\ data is the representation
of a likelihood function, whereas our ouptut is a
representation of a posterior pdf, one per star.
Posterior information used naively can lead to serious statistical
errors, because each data point has had a prior multiplied in.
Multiplicative uses of the data will effectively take the prior
pdf to a large power.
We have (in previous work)
given examples of correct uses of posterior samplings or pdfs
for subsequent inferences (\citealt{hogg08}; \citealt{dfm14}); we
encourage power-users to consult those methodological
contributions before engaging with these outputs.

@@

\subsection{What's That Feature?}
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{whatsThatFeature.pdf}
\caption{Features in the \cmd: The expectation values from the posterior pdfs of the true parallax}
\label{fig:wtf}
\end{figure}

The denoised expectation values of the parallax projected into the \cmd\ show interest features.
We visualize this \cmd\ again in Figure \ref{fig:wtf} pointing out some interesting features.
\begin{itemize}
\item (1) Narrow Lower Main Sequence
\item (2) Wide Upper Main Sequence
\item (3) Plausible Binary Sequence
\item (4) Prominent Red clump
\item (5) Narrow Red Giant Branch
\item (6) Noticeable turn off
\end{itemize}
The lower main sequence is very narrow while the upper main sequence is much wider. This may be an age variation on the main sequence mixed with a malmquist bias of seeing a much larger volume of upper main sequence stars. There is a plausible binary sequence of stars with about twice the brightness as the main sequence. The helium burning red clump is (surprisingly) prominent for the volume DR1 probed, and the red giant branch is fairly straight and narrow. The main sequence turn off is noticeable and also surprisingly narrow, possibly a reflection of the star formation history of the local volume also mixed with a malmquist bias.

@@

In Section~\ref{sec:method} we listed a set of assumptions upon which
the photometric parallax method in this project is based.
Here we return to these assumptions, and discuss them critically in more
detail.
\begin{description}
\item[stationarity] (DWH needs to clean this paragraph)
  We have assumed that (locally) the morphology of
  the \cmd\ is independent of signal-to-noise; that is, we have assumed that
  both the low and high
  signal-to-noise objects are being drawn from the same distribution.
  This isn't precisely true because nearby stars in \tgas\ are
  mostly main sequence stars, while we can see luminous giants to much larger
  distances. So the red giant branch is a population of
  relatively lower signal to noise objects compared with the main
  sequence. This does not violate the stationarity assumption in
  detail, because the low and high signal to noise objects have to be
  drawn from the same distribution within some feature or patch of the
  space. This is all assuming the Gaia noise model is correct. If the
  uncertainties are over estimated then we could be over deconvolving
  (generating a tighter distribution than the truth) And in some
  extreme over estimate of uncertainties the lower signal to noise
  giants could be pulled to the main sequence. But with the quoted
  uncertainties this seems unlikely. As you move up the giant branch,
  and probe farther giants, the signal to noise will continue to
  diminish, but because the color uncertainties are not large, the S/N
  spread within a patch of the \cmd\ is well mixed. Possibly a
  metallicity trend with distance, placing more nearby (higher S/N)
  objects at slightly different positions on the CMD pulling lower S/N
  objects to that distribution for our prior. But if something has
  small error bars it is good data is good data both for inference and
  generating the prior.

  For very rare kinds of stars. Inference with our prior is bad for
  underrepresented populations in the \cmd. For example, it appears
  there are only $2$ white dwarfs in \gaia\ DR1. With them being very
  underrepresented in the prior, the amplitudes of the gaussians
  representing them is very small, so the WDs are more likely to be
  pulled to mimic a much farther away MS star.
\item[selection] DWH
\item[big data] DWH
\item[noise model] DWH
\item[dust] Within our probabilistic framework, we improperly handle
  dust. We are taking a point estimate of the probabilistic dust model
  at a quantile of our distance posterior, correcting our photometry
  using this point estimate, and then re-deriving the prior. A full,
  proper account would sample the dust model at distances sampled from
  the posterior, and propagate those samples back into the prior
  during the iterative process. The reason we don't do a full
  probabilistic treatment, is that this proper dust distribution would
  not be a Gaussian distribution, which \xd\ requires. We could
  calculate this more complex dust distribution and estimate it as a
  Gaussian distribution to feed to \xd, however, this is beyond the
  scope of this paper.
\item[mixture of Gaussians] DWH
\item[no physics] DWH
\end{description}

@@

\acknowledgments It is a pleasure to thank
  Ana Bonaca (Harvard),
  Andy Casey (Monash),
  Stephen Feeney (Flatiron),
  Dustin Lang (Toronto),
  Jo Bovy (Toronto)
and the attendees at the Stars Group Meeting at the Flatiron Institute
Center for Computational Astrophysics for comments and input.
This project was developed in part at the 2016 \acronym{NYC} Gaia Sprint, hosted
by the Center for Computational Astrophysics at the Simons Foundation
in New York City.

>>>>

Under these assumptions, we would like to infer the true parallax, $\varpi_{\true}$, for each star given its measured parallax, $\varpi$, and associated uncertainty, $\sigma_{\varpi}$, from \tgas.
A first application of Bayes theorem leads to
\begin{equation}
p(\varpi_{\true} \given \varpi, \sigma^2_{\varpi}) = \frac{1}{Z}\,p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi}) \, p(\varpi_{\true}) \quad ,
\label{eq:bayes}
\end{equation}
where $p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi})$ is the likelihood of the observed parallax, $p(\varpi_{\true})$ is a prior pdf for the true
parallax, and $Z$ is a normalization constant (the model evidence, or marginalized likelihood).
The strength of our inference lies in how we generate our prior.
Instead of assuming a functional form for the spatial distribution of stars in the Milky Way (\citealt{astraatmadja16b}), or relying on stellar models (\citealt{gaia16}), we build a noise-deconvolved estimate of the \cmd\ from the observed colors and absolute magnitudes of our full set of stars.
This is because supplementary to the trigonometric measure of distance, the photometry of stars is also informative of their distances.
The intrinsic temperatures and luminosities of stars tightly correlate leading to the observed colors and absolute magnitudes of stars also tightly correlating, usually visualized as the \cmd.
The tight correlation of the \cmd\ benefits our inference because the luminosity or absolute magnitude of a star $M$, when compared with its brightness or apparent magnitude $m$, contains distance information $d$ via the fundamental relation $m=M+5\log(d/10\, \mathrm{pc})$.
We include this in the inference by using the \cmd\ as part of our prior for $\varpi_{\true}$.
Specifically, we use the \cmd\ generated from the data themselves, as detailed in the next section.

@@

We can now write the multidimensional analogous to Equation \ref{eq:bayes},
\begin{equation}
p(\mathbf{Y}_n \given \mathbf{y}_n, I_n) = p(\mathbf{y}_n \given \mathbf{Y}_n, I_n) \, p(\mathbf{Y}_n) \label{eq:posterior}
\end{equation}
where $p(\mathbf{y}_n \given \mathbf{Y}_n, I_n)$ is the likelihood of the data vector, and $p(\mathbf{Y}_n)$ is our prior.
Before we describe how this can be turned into a posterior belief on the true parallax, we turn our attention to the prior, which will be generated from the data themselves. We will use XD, which will take advantage of the Gaussian likelihood.

@@

To infer $\varpi_{\true}$ once the \cmd\ prior is constructed, we first do the inference in the 2D space of the \cmd\ prior $p(\mathbf{Y} \given \alpha)$, using Equation \ref{eq:posterior}. We then project the 2D posterior into the 1D transformed absolute magnitude space, and do a little algebra to solve for $\varpi_{\true}$

@@

@@

\BLcomment{Isn't this repeating what was described in the previous paragraph?}
We need to correct the photometry used in our method, both for the generation of the prior and the likelihood, for dust. With the emergence of 3D dust maps, it is now possible to estimate dust corrections to stars within the Milky Way. The challenge is getting a proper estimate of the distance to the star before inferring $\varpi_{\true}$, since we need a dust correction prior to doing our inference. Dust corrections are particularly non trivial for stars with poor parallax signal to noise. Stars with poor signal to noise have a likelihood that is consistent with infinite distance and can therefore have severe dust corrections. This is most obvious in the giant stars, which tend to have the lowest signal to noise, and can therefore have dust corrections $> 2$ magnitudes. Although the dust model is probabilistic, as well as our distance inference, to move forward, we break from a fully Bayesian framework, in which we might infer the distance and the dust simultaneously (more about this in the discussion). Instead, we correct for dust using a point estimate from the 3D dust map \citep{green15} sampled at a distance estimated from iteratively inferring and then sampling our posterior pdf over $\varpi_{\true}$. In more detail, to apply a dust correction to the \tmass\ photometry, we first generate our prior using the observed, attenuated photometry. Using this raw prior, we infer more precise parallaxes to all the stars in our sample. We then use this more precise parallax posterior to get a measurement of dust in the 3D dust map for each star. And repeat. This approach minimizes severe distances, and their associated severe dust corrections, which would occur if using the likelihoods alone. In detail, we take the $5\%$ quantile of each distance posterior, so the closest bit of the posterior, and sample the 3D dust map at that distance. We take the $5\%$ quantile, as apposed to the median, for example, to do a light dust correction, because over correcting for dust can shoot the likelihood into an unphysical, different space of the \cmd\ and bias the parallax inference. For each distance, we sample the median of the probabilistic dust model. We apply each light dust correction to our \tmass\ photometry given by Equation \ref{eq:data} where $Q_{\lambda}$ = [0.709, 0.302] for bands [J, K] respectively, taken from table (LMA REF TABLE SFD). We then regenerate our prior using the lightly dust corrected photometry. The dust values we obtain in our final iteration, and which we apply to our photometry, are visualized in galactic coordinates in Figure \ref{fig:dust}. After applying a dust correction, we do not update the uncertainties, but keep them the same as before the dust correction, which is wrong and under representing our uncertainties but this leads, again, to a lighter deconvolution. \BLedit{In other words, the features of our inferred \cmd\ are conservatively broad (broader than if we over-estimated the errors which would have lead to excessive deconvolution and unrealistically narrow features).} We could add the covariance of the dust to our likelihoods, but this is beyond the scope of this demonstration of concept. We iterate this process 10 times, however the dust values seem to converge after about 2-3 iterations. The iterative process allows for slight variations in distance and corresponding dust to become more consistent with the \cmd.

@@

\subsection{Shrinkage of parallax uncertainties}

@@

\BLcomment{Need to discuss the implications of that.}

@@

We obtained the photometric parallaxes in this project
by building a data-driven model of \cmd\ of the stars in the \tgas\ data set,
and using it as a prior pdf for Bayesian inference.
The posterior pdfs for distance that we obtain are, in general, much
narrower than the likelihood functions delivered by the
\gaia\ Mission, and therefore the distance estimates (or,
equivalently, parallax estimates) are much more precise.
It is not surprising that a Bayesian inference provides more
precise inferences than the likelihood function alone; Bayesian
inferences bring in new information that decrease variance (but
can introduce bias).
We have shown that, in addition to pure precision improvements, at
least some aspects of \emph{accuracy} have been improved as well:
The posterior distance estimates to stars in (at least one) stellar
cluster are much more clustered than likelihood-based distance
estimates.

Although we have not performed principled hierarchical Bayesian inference in this
project, we built our prior for each star's parallax by
performing a statistically responsible deconvolution
of the \gaia\ \tgas\ data that is justifiable under a clear set of
assumptions.
This deconvolution, using a reasonable noise model and an
assumption of stationarity, shrinks the parallax uncertainties, most dramatically for the stars measured at lower signal-to-noise.
Because the method we use, \xd, which is
an Empirical Bayesian maximum-marginalized-likelihood estimator, accounts for
heteroskedastic noise, we were able to build our photometric parallax
model---which is a model of the \cmd---using
all the data, not just the data with the highest signal-to-noise.
So, the model for the \cmd\ we built is representative for
our subset of the \tgas\ Catalog.
Because our model is representitive of our full subset catalog, we expect any (sensible) parallax estimates
we generate from our parallax posterior pdfs to be only weakly biased.

@@

\subsection{What's That Feature?}
The denoised expectation values of the parallax projected into the \cmd\ show interest features.
We visualize this \cmd\ again in Figure \ref{fig:wtf} pointing out some interesting features.
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{whatsThatFeature.pdf}
\caption{Features in the \cmd: The expectation values from the posterior pdfs of the true parallax}
\label{fig:wtf}
\end{figure}

The lower main sequence is very narrow while the upper main sequence
is much wider. This may be an age variation on the main sequence mixed
with a malmquist bias of seeing a much larger volume of upper main
sequence stars. There is a plausible binary sequence of stars with
about twice the brightness as the main sequence. The helium burning
red clump is prominent for the volume DR1 probed, and the red giant
branch is fairly straight and narrow. The main sequence turn off is
noticeable and also surprisingly narrow, possibly a reflection of the
star formation history of the local volume also mixed with a malmquist
bias.

@@

In Section~\ref{sec:method} we listed a set of assumptions upon which
the photometric parallax method in this project is based.
Here we return to these assumptions, and discuss them critically in more
detail, under the same set of assumption labels:
\begin{description}
\item[stationarity] (DWH needs to clean this paragraph)
  We have assumed that (locally) the morphology of
  the \cmd\ is independent of signal-to-noise; that is, we have assumed that
  both the low and high
  signal-to-noise objects are being drawn from the same distribution.
  This isn't precisely true because nearby stars in \tgas\ are
  mostly main sequence stars, while we can see luminous giants to much larger
  distances. So the red giant branch is a population of
  relatively lower signal to noise objects compared with the main
  sequence. This does not violate the stationarity assumption in
  detail, because the low and high signal to noise objects have to be
  drawn from the same distribution within some feature or patch of the
  \cmd\ space.
\item[selection] The biggest limitation of our method for generating
  the \cmd\ from the data is that we made no use of any kind of
  \tgas\ selection function or completeness estimate or
  inverse-selection-volume corrections.  This model is a model of the
  \tgas\ Catalog (or really the \tgas--\tmass\ intersection), not of
  the stars in the Milky Way, nor any volume-limited subsample of the
  Milky Way.  Importantly, because we used \emph{all} of the stars in
  the \tgas--\tmass\ intersection, the model \emph{is} representive of
  the full intersection, even the distant parts, where all the
  parallaxes are measured at low signal-to-noise.  This is in contrast
  to techniques that might build the model from only the high
  signal-to-noise sources; such a model would be biased towards
  stellar types and compositions found locally, as well as
  lower-luminosity stars, and stars which happened to get good
  parallaxes.  However, our use of the entire
  \tgas--\tmass\ intersection without any selection function restricts
  the use of our \cmd\ model as a prior to inferences within that same
  \tgas--\tmass\ intersection.  It would be biasing to use this same
  prior to infer distances for the full billion-star catalog released
  in \gaia\ DR1 alongside \tgas, or for stars fainter or brighter or
  bluer than those included in the \tgas--\tmass\ intersection.
\item[big data] The empirical Bayesian methodology is a good
  approximation to full hierarchical inference in the limit of large
  numbers of data points, such that no individual star is carrying a
  lot of weight in the model of the \cmd.  With its millions of stars,
  the \tgas\ Catalog appears to safely live in this limit.  However,
  because there are parts of the \cmd\ that are poorly populated, even
  in \tgas, the empirical Bayes approximation may be bad for some
  portions of the \cmd.  In particular, there appear to be essentially
  no white dwarf stars, and other classes (like bright red giant
  branch stars) have few to no exemplars in the data set at high
  signal-to-noise (in parallax).  For all these reasons, we do think
  that this approximation is not perfectly safe, and it is a goal to
  explore full hierarchical modeling in the coming years (and see, for
  example, \citealt{leistedtHogg2017}).
\item[noise model] Everything in both the establishment of the data-driven
  prior and the use of it for parallax inference has been done under the deep
  assumption that the \gaia\ noise model is correct. If the
  uncertainties are over-estimated then we could be over-deconvolving.
  That is, the data-driven model could be producing a tighter or more
  informative distribution than what would truly be observed in a much
  higher signal-to-noise experiment or survey. That is, incorrect uncertainty
  estimates will bias the empirical-Bayes prior we obtain. On the inference
  side, wrong uncertainty estimates lead to further biases in the derived photometric parallaxes.
\item[dust] Within our probabilistic framework, we improperly handle
  dust. We are taking a point estimate of the probabilistic dust model
  at a quantile of our distance posterior, correcting our photometry
  using this point estimate, and then re-deriving the prior. A full,
  proper account would sample the dust model at distances sampled from
  the posterior, and propagate those samples back into the prior
  during the iterative process. The reason we don't do a full
  probabilistic treatment, is that this proper dust distribution would
  not be a Gaussian distribution, which \xd\ requires. We could
  calculate this more complex dust distribution and estimate it as a
  Gaussian distribution to feed to \xd, however, this is beyond the
  scope of this paper.
\item[mixture of Gaussians] In everything in this project DWH
\item[no physics] DWH
\end{description}

The outputs of this project include posterior pdfs for stellar parallaxes.
It is tempting to treat these outputs as equivalent to some kind of catalog of
measurements, as we treated the original parallax measurements from \tgas.
That is, it is tempting to treat these as simply ``better'' measurements.
Taken one star at a time, this is permitted and true, and the basis for our
claim that we have de-noised the \tgas\ Catalog.
However, there are important differences, as there are with all
probabilistic catalogs (\citealt{hogg11}; \citealt{portillo17}), between the
original \tgas\ measurements and the de-noised measurements.
For users of our output, the most important difference between our
output and the \tgas\ data input is that the \tgas\ data is the representation
of a likelihood function, whereas our ouptut is a
representation of a posterior pdf, one per star.
Posterior information used naively can lead to serious statistical
errors, because each data point has had a prior multiplied in.
Multiplicative uses of the data will effectively take the prior
pdf to a large power.
We have (in previous work)
given examples of correct uses of posterior samplings or pdfs
for subsequent inferences (\citealt{hogg08}; \citealt{dfm14}); we
encourage power-users to consult those methodological
contributions before engaging with these outputs.

@@

\acknowledgments It is a pleasure to thank
  Ana Bonaca (Harvard),
  Jo Bovy (Toronto),
  Andy Casey (Monash),
  Stephen Feeney (Flatiron),
  Dustin Lang (Toronto),
  David Spergel (Flatiron),
and the attendees at the Stars Group Meeting at the Flatiron Institute
Center for Computational Astrophysics for comments and input.
This project was developed in part at the 2016 \acronym{NYC} Gaia Sprint, hosted
by the Center for Computational Astrophysics at the Simons Foundation
in New York City.
